---
title: "213 Data Analysis"
output: html_notebook
---

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(ez)
```

# Behavioral Data

```{r}
behavioral.data <- read_csv('data/behavioral_data_tidy.csv')
```

The data has three columns

```{r}
behavioral.data
```

Let's summarize the behavioral data to get means and SDs for each condition.

```{r}
behavioral.data <- behavioral.data %>% mutate(PercentCorrect = (CorrectResponses / 29)*100)
behavioral.data %>% 
  group_by(Condition) %>% 
  summarize(M=mean(PercentCorrect), SD = sd(PercentCorrect))
```

Let's run two sets of t-tests, to see if there is a difference in accuracy between Literal and Metaphorical and/or Literal and Anomalous.

```{r}
t.test.data.literal.metaphorical <- behavioral.data %>% filter(Condition %in% c('Literal', 'Metaphorical'))
t.test(CorrectResponses ~ Condition, data=t.test.data.literal.metaphorical, paired=T)
```

```{r}
t.test.data.literal.anomalous <- behavioral.data %>% filter(Condition %in% c('Literal', 'Anomalous'))
t.test(CorrectResponses ~ Condition, data=t.test.data.literal.anomalous, paired=T)
```

# Graphing Behavioral Data

```{r}
ggplot(behavioral.data, aes(x=Condition, y=PercentCorrect))+
  geom_boxplot()
```

# Loading EEG Data

We will start by loading in our data.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
eeg.data <- read_csv('data/voltage_data_tidy.csv')
```

The data has five columns. 

* **t** represents the time in milliseconds relative to the critical word.
* **electrode** is the number of the electrode on the EGI 128-electrode cap.
* **voltage** is measured in microvolts.
* **subject** is an anonymous subject code.
* **condition** is one of three conditions: Anomalous, Literal, or Metaphorical.

A subset of the data is shown below.

```{r echo=FALSE}
eeg.data[1:1000, ]
```

# Grand Average Waveforms

Our analysis will be on three midline electrodes, Fz (EGI #11), Pz (EGI #62), and Cz (EGI #129). The target time window for the expected N400 effect is 300 to 500ms.

```{r}
target.electrodes <- c(11,62,129)
n400.window <- 300:500
```

Extracting data...

```{r message=FALSE, warning=FALSE}
library(dplyr)

grand.average.data <- eeg.data %>% filter(electrode %in% target.electrodes) %>% 
  group_by(t, condition, electrode) %>% summarize(mean.voltage = mean(voltage))
```

Plotting data...

```{r}
library(ggplot2)

ggplot(grand.average.data, aes(x=t, y=mean.voltage, color=condition))+
  facet_wrap(~electrode)+
  geom_vline(xintercept = 0)+
  geom_hline(yintercept = 0)+
  annotate("rect", fill="black", alpha=0.1, xmin=min(n400.window),xmax=max(n400.window), ymin=-5, ymax=5)+
  labs(x="Time relative to stimulus onset (ms)", y="Mean voltage (microvolts)")+
  geom_line()+
  theme_minimal()
```


# Statistical Analysis

Gathering data for the ANOVA.

```{r}
anova.data <- eeg.data %>% filter(electrode %in% target.electrodes, t %in% n400.window) %>%
  group_by(subject, electrode, condition) %>% summarize(mean.voltage = mean(voltage))

anova.data$electrode <- as.factor(anova.data$electrode)
```

Computing ANOVA.

```{r message=FALSE, warning=FALSE}
result <- ezANOVA(anova.data, dv=mean.voltage, wid=subject, within=c(electrode, condition))

result$ANOVA
```