---
title: "An ERP study on metaphorical language processing"
author: "Josh de Leeuw"
format: html
editor: visual
---

```{r}
#| echo: false
#| label: Load R libraries
#| warning: false

library(readr)
library(dplyr)
library(ggplot2)
library(ez)

```

## Introduction

This is a study about metaphor using ERP methods.

## Methods

```{r}
#| echo: false
#| message: false
#| label: Load behavioral data

behavioral.data <- read_csv('data/behavioral_data_tidy.csv')

```

```{r}
#| echo: false
#| label: Count number of subjects

n.subjects <- behavioral.data %>% 
  group_by(Subject) %>% 
  tally() %>% 
  nrow()
```

### Participants

`r n.subjects` Vassar students participated in this study.

### Procedure

Read sentences. Record brainwaves.

### EEG Recordings

Put electrodes on head. Voltages go brrrrrrr.

## Results

### Behavior

```{r}
#| label: Summarize behavioral data
#| message: false
#| echo: false

behavioral.summary <- behavioral.data %>%
  mutate(PercentCorrect = CorrectResponses/29 * 100) %>%
  group_by(Condition) %>%
  summarize(M = mean(PercentCorrect), SD = sd(PercentCorrect))

```

We checked for comprehension accuracy of the sentences by looking at the percent of "sensible" responses for each sentence type. Participants said the sentences made sense more in the literal (*M* = `r behavioral.summary %>% filter(Condition == "Literal") %>% pull(M) %>% round(2)`%, *SD* = `r behavioral.summary %>% filter(Condition == "Literal") %>% pull(SD) %>% round(2)`%) and anomalous (*M* = `r behavioral.summary %>% filter(Condition == "Anomalous") %>% pull(M) %>% round(2)`%, *SD* = `r behavioral.summary %>% filter(Condition == "Anomalous") %>% pull(SD) %>% round(2)`%) conditions than in the metaphorical condition (*M* = `r behavioral.summary %>% filter(Condition == "Metaphorical") %>% pull(M) %>% round(2)`%, *SD* = `r behavioral.summary %>% filter(Condition == "Metaphorical") %>% pull(SD) %>% round(2)`%), suggesting that the metaphorical meanings were slightly more difficult to understand (see @fig-behavioral-accuracy ).

```{r}
#| label: fig-behavioral-accuracy
#| fig-cap: The number of correct responses to the comprehension questions was slightly lower for metaphorical sentences.
#| echo: false

ggplot(behavioral.data, aes(x=Condition, y=CorrectResponses, color=Condition))+
  geom_jitter(width=0.25, height = 0, size=2)+
  scale_color_hue(guide="none")+
  labs(x="Sentence Type", y="Number of Correct Responses\n(out of 29)")+
  theme_bw()
```

### ERP

We measured the amplitude of the ERP in the 300-500ms window at four electrode sites: F3, Fz, F4, and Cz. @fig-erp shows the grand average wave forms.

```{r}
#| label: Load EEG data
#| echo: false
#| message: false

eeg.data <- read_csv('data/voltage_data_tidy.csv')
```

```{r}
#| label: Define ERP electrodes and window
#| echo: false

target.electrodes <- c(11,24,124,129)
target.window <- 300:500
```

```{r}
#| label: Calculate grand average ERPs
#| echo: false
#| message: false

grand.average.erp <- eeg.data %>%
  group_by(t, electrode, condition) %>%
  filter(electrode %in% target.electrodes) %>%
  summarize(m = mean(voltage), se = sd(voltage)/sqrt(n()))
```

```{r}
#| label: fig-erp
#| fig-cap: Grand Average ERPs for the four target electrodes (Fz, F3, F4, Cz). The shaded window represents the target time window for our analysis of the N4 ERP component.
#| echo: false

grand.average.erp <- grand.average.erp %>%
  ungroup() %>%
  mutate(electrode.name = case_when(
    electrode == 11 ~ "Fz",
    electrode == 24 ~ "F3",
    electrode == 124 ~ "F4",
    electrode == 129 ~ "Cz",
    .default = as.character(electrode)
  ))

ggplot(grand.average.erp, aes(x=t, y=m, ymin=m-se, ymax=m+se, color=condition, fill=condition))+
  facet_wrap(vars(electrode.name))+
  geom_vline(xintercept = 0)+
  geom_hline(yintercept = 0)+
  annotate("rect", fill="black", alpha=0.1, xmin=min(target.window),xmax=max(target.window), ymin=-5, ymax=5)+
  labs(x="Time relative to stimulus onset (ms)", y="Mean voltage (microvolts)")+
  geom_line()+
  geom_ribbon(color=NA, alpha=0.3)+
  scale_y_reverse()+
  theme_minimal()+
  theme(panel.grid = element_blank())

```

#### Is there a difference in N400 for different sentence types?

```{r}
#| label: calculate anova data
#| echo: false
#| message: false
erp.anova.data <- eeg.data %>%
  filter(t %in% target.window, electrode %in% target.electrodes) %>%
  group_by(subject, condition, electrode) %>%
  mutate(electrode = factor(electrode),
         subject = factor(subject),
         condition = factor(condition)) %>%
  summarize(mean.subject.amplitude = mean(voltage))
```

```{r}
#| label: run ANOVA
#| echo: false

erp.anova <- ezANOVA(erp.anova.data, dv=mean.subject.amplitude, within=c(condition, electrode), wid=subject)

erp.anova$ANOVA
```
